# Intelligent Fitness Tracking: Calorie Prediction & Pattern Discovery

##  Problem Statement
Accurate calorie estimation is vital for fitness tracking apps. This project utilizes physiological signals (Heart Rate, BMI, Duration) and contextual data to:
1. **Predict** calories burned using Regression models.
2. **Discover** hidden workout intensity clusters using Unsupervised Learning.

##  Key Features
- **Supervised Learning:** Compares Linear Regression, KNN, Random Forest, and XGBoost to achieve R² ≥ 0.80.
- **Unsupervised Learning:** Uses PCA for feature compression and KMeans for user segmentation (Target Silhouette ≥ 0.15).
- **Automated Preprocessing:** Handles outliers via IQR and automates feature scaling/encoding.

## Tech Stack
- **Languages:** Python 3.x
- **Libraries:** Pandas, NumPy, Scikit-Learn, XGBoost, Matplotlib, Seaborn
- **Validation:** 5-Fold Cross-Validation & Fixed Random Seeds (Seed 42)

##  Results Summary
- **Regression:** Achieved high precision for real-time calorie estimation.
- **Clustering:** Identified 3 distinct physiological intensity profiles.
"This project creates an intelligent fitness assistant that doesn't just guess calories, but actually understands a user's unique physiological response to exercise, allowing for personalized coaching and safer workout recommendations."


Key Highlights:
Supervised Learning: 
	Implemented XGBoost and KNN Regressors to predict calorie expenditure with a target accuracy of $R^2 \geq 0.80$.
Unsupervised Learning: 
	Applied PCA for feature compression and KMeans Clustering to identify hidden workout intensity segments (Target Silhouette $\geq 0.15$).
Data Pipeline: 
	Automated preprocessing using IQR Outlier detection, One-Hot Encoding, and Standard Scaling.
Validation:
	Ensured strict reproducibility with Fixed Random Seeds and Cross-Validation.
Business Goal: 
	Providing real-time energy expenditure insights and personalized fitness coaching archetypes.